---
title: "Drop Count Analysis"
author: "Jimmy Larson"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
## load packages----
library(tidyverse)
library(ggpubr)
library(car)
library(lubridate)
library(RColorBrewer)
#install.packages("ggrepel")
library(ggrepel)
#install.packages("hrbrthemes")
library(hrbrthemes)
## plotting theme----
theme_jl <- theme_ipsum_ps(grid="XY", axis="xy", axis_text_size = 10, axis_title_size = 11, axis_col = "black")
## load data----
dropAug25 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_aug_25.csv")
dropSep1 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_sep_1.csv")
dropSep8 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_sep_8.csv")
dropSep15 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_sep_15.csv")
dropSep22 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_sep_22.csv")
dropSep29 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_sep_29.csv")
dropOct7 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_oct_7.csv")
dropOct13 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_oct_13.csv")
dropOct21 <- read_csv("data/2020/dropCounts/pfd_fruit_drop_counts_oct_21.csv")
treeCounts <- read_csv("data/2020/dropCounts/pfd_fruit_tree_counts_oct_21.csv")
## calculate remaining fruit on treee----
treeCounts %>%
  rowwise() %>%
  mutate(remainingFruit = sum(count1,count2,count3,count4,count5,count6,count7, na.rm = TRUE)) %>%
  select(row, tree, plot, treatment, rep, flag, remainingFruit) %>%
  left_join(dropAug25) %>%
  rename(Aug25Drop = dropped) %>%
  left_join(dropSep1) %>%
  rename(Sep1Drop = dropped)%>%
  mutate(cumDropSep1 = Aug25Drop + Sep1Drop) %>%
  left_join(dropSep8) %>%
  rename(Sep8Drop = dropped)%>%
  mutate(cumDropSep8 = Sep8Drop + cumDropSep1) %>%
  left_join(dropSep15) %>%
  rename(Sep15Drop = dropped)%>%
  mutate(cumDropSep15 = Sep15Drop + cumDropSep8) %>%
  left_join(dropSep22) %>%
  rename(Sep22Drop = dropped)%>%
  mutate(cumDropSep22 = Sep22Drop + cumDropSep15) %>%
  left_join(dropSep29) %>%
  rename(Sep29Drop = dropped)%>%
  mutate(cumDropSep29 = Sep29Drop + cumDropSep22) %>%
  left_join(dropOct7) %>%
  rename(Oct7Drop = dropped)%>%
  mutate(cumDropOct7 = Oct7Drop + cumDropSep29) %>%
  left_join(dropOct13) %>%
  rename(Oct13Drop = dropped)%>%
  mutate(cumDropOct13 = Oct13Drop + cumDropOct7) %>%
  left_join(dropOct21) %>%
  rename(Oct21Drop = dropped) %>%
  mutate(cumDropOct21 = Oct21Drop + cumDropOct13) %>%
  mutate(totalFruit = cumDropOct21 + remainingFruit) %>%
  mutate(perDropAug25 = ((Aug25Drop / totalFruit)*100),
         perDropSep1 = ((cumDropSep1 / totalFruit)*100),
         perDropSep8 = ((cumDropSep8 / totalFruit)*100),
         perDropSep15 = ((cumDropSep15 / totalFruit)*100),
         perDropSep22 = ((cumDropSep22 / totalFruit)*100),
         perDropSep29 = ((cumDropSep29 / totalFruit)*100),
         perDropOct7 = ((cumDropOct7 / totalFruit)*100),
         perDropOct13 = ((cumDropOct13 / totalFruit)*100),
         perDropOct21 = ((cumDropOct21 / totalFruit)*100))%>%
  select(row, tree, plot, treatment, rep, perDropAug25, perDropSep1, perDropSep8, perDropSep15, perDropSep22,
         perDropSep29, perDropOct7, perDropOct13, perDropOct21)-> drops
## create long form data ----
drops %>%
  gather('perDropAug25', 'perDropSep1', 'perDropSep8', 'perDropSep15', 'perDropSep22', 'perDropSep29',
         'perDropOct7', 'perDropOct13', 'perDropOct21', key = "measDate", value = "drops") %>%
  mutate(date = case_when(measDate == 'perDropAug25' ~ mdy("8/25/2020"), ### add date
                   measDate == 'perDropSep1' ~ mdy("9/1/2020"),
                   measDate == 'perDropSep8' ~ mdy("9/8/2020"),
                   measDate == 'perDropSep15' ~ mdy("9/15/2020"),
                   measDate == 'perDropSep22' ~ mdy("9/22/2020"),
                   measDate == 'perDropSep29' ~ mdy("9/29/2020"),
                   measDate == 'perDropOct7' ~ mdy("10/7/2020"),
                   measDate == 'perDropOct13' ~ mdy("10/13/2020"),
                   measDate == 'perDropOct21' ~ mdy("10/21/2020"))) %>%
  select(row, tree, plot, treatment, rep, drops, date) -> dropsLong
## explore data ----
ggplot(dropsLong, aes(x = treatment, y = drops, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~ date, scales = "free") +
  scale_fill_brewer(palette = "RdYlBu") +
  labs(y = "Percent Fruit Drop (%)",
       x = "Treatment") +
  theme_jl
## Analyze and plot drop percent----
### anova test
drops.aov <- aov(drops ~ treatment*date, data = dropsLong)
summary.aov(drops.aov) # treatment*date interaction is significant, analyze within dates
### anova for each date
#### Aug 25----
dropAug25.aov <- aov(perDropAug25 ~ treatment, data = drops)
summary.aov(dropAug25.aov)
Aug25Model <- data.frame(fit = fitted(dropAug25.aov),
                         res = resid(dropAug25.aov),
                         treat = drops$treatment)
#### test normality
Aug25QQ <- ggplot(drops, aes(sample = perDropAug25)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Aug. 25") +
  theme_jl
dropAug25.resid <- residuals(object = dropAug25.aov)
#shapiro.test(x = dropAug25.resid)## p < 0.05
shapiro.test(x = sqrt(dropAug25.resid)) ## p > 0.05
##### test homogeniety of variance
Aug25Resid <- ggplot(Aug25Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values", 
       subtitle = "Residual Plot-Aug. 25")+
  theme_jl
leveneTest(perDropAug25 ~ treatment, data = drops) ## p > 0.05
#### mean separation
dropAug25.norm.aov <- aov(sqrt(perDropAug25) ~ treatment, data = drops) # had to transform data with sqrt to normalize
TukeyHSD(dropAug25.norm.aov)
#### Sep 1----
dropSep1.aov <- aov(perDropSep1 ~ treatment, data = drops)
summary.aov(dropSep1.aov)
Sep1Model <- data.frame(fit = fitted(dropSep1.aov),
                         res = resid(dropSep1.aov),
                         treat = drops$treatment)
##### test normality
Sep1QQ <- ggplot(drops, aes(sample = perDropSep1)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Sep. 1") +
  theme_jl
dropSep1.resid <- residuals(object = dropSep1.aov)
shapiro.test(x = dropSep1.resid) ## p > 0.05
##### test homogeniety of variance
Sep1Resid <- ggplot(Sep1Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Sep. 1")+
  theme_jl
leveneTest(perDropSep1 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropSep1.aov)
#### Sep 8----
dropSep8.aov <- aov(perDropSep8 ~ treatment, data = drops)
summary.aov(dropSep8.aov)
Sep8Model <- data.frame(fit = fitted(dropSep8.aov),
                        res = resid(dropSep8.aov),
                        treat = drops$treatment)
#### test normality
Sep8QQ <- ggplot(drops, aes(sample = perDropSep8)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Sep. 8") +
  theme_jl
dropSep8.resid <- residuals(object = dropSep8.aov)
shapiro.test(x = dropSep8.resid) ## p > 0.05
##### test homogeniety of variance
Sep8Resid <- ggplot(Sep8Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Sep. 8")+
  theme_jl
leveneTest(perDropSep8 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropSep8.aov)
#### Sep 15----
dropSep15.aov <- aov(perDropSep15 ~ treatment, data = drops)
summary.aov(dropSep15.aov)
Sep15Model <- data.frame(fit = fitted(dropSep15.aov),
                        res = resid(dropSep15.aov),
                        treat = drops$treatment)
#### test normality
Sep15QQ <- ggplot(drops, aes(sample = perDropSep15)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Sep. 15") +
  theme_jl
dropSep15.resid <- residuals(object = dropSep15.aov)
shapiro.test(x = dropSep15.resid)## p > 0.05
##### test homogeniety of variance
Sep15Resid <- ggplot(Sep15Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Sep. 15")+
  theme_jl
leveneTest(perDropSep15 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropSep15.aov)
#### Sep 22----
dropSep22.aov <- aov(perDropSep22 ~ treatment, data = drops)
summary.aov(dropSep22.aov)
Sep22Model <- data.frame(fit = fitted(dropSep22.aov),
                        res = resid(dropSep22.aov),
                        treat = drops$treatment)
#### test normality
Sep22QQ <- ggplot(drops, aes(sample = perDropSep22)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Sep. 22") +
  theme_jl
dropSep22.resid <- residuals(object = dropSep22.aov)
#shapiro.test(x = dropSep22.resid) ## p<0.05
shapiro.test(x = (dropSep22.resid)^(-.5)) ## p>0.05
##### test homogeniety of variance
Sep22Resid <- ggplot(Sep22Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Sep. 22")+
  theme_jl
leveneTest(perDropSep22 ~ treatment, data = drops) ## p > 0.05
#### mean separation
dropSep22.norm.aov <- aov((perDropSep22)^(-.5) ~ treatment, data = drops) ## used 1/sqrt() to normalize
TukeyHSD(dropSep22.norm.aov)
#### Sep 29----
dropSep29.aov <- aov(perDropSep29 ~ treatment, data = drops)
summary.aov(dropSep29.aov)
Sep29Model <- data.frame(fit = fitted(dropSep29.aov),
                        res = resid(dropSep29.aov),
                        treat = drops$treatment)
#### test normality
Sep29QQ <- ggplot(drops, aes(sample = perDropSep29)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Sep. 29") +
  theme_jl
dropSep29.resid <- residuals(object = dropSep29.aov)
shapiro.test(x = dropSep29.resid) ## p>0.05
##### test homogeniety of variance
Sep29Resid <- ggplot(Sep29Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Sep. 29")+
  theme_jl
leveneTest(perDropSep29 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropSep29.aov)
#### Oct 7----
dropOct7.aov <- aov(perDropOct7 ~ treatment, data = drops)
summary.aov(dropOct7.aov)
Oct7Model <- data.frame(fit = fitted(dropOct7.aov),
                         res = resid(dropOct7.aov),
                         treat = drops$treatment)
#### test normality
Oct7QQ <- ggplot(drops, aes(sample = perDropOct7)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Oct. 7") +
  theme_jl
dropOct7.resid <- residuals(object = dropOct7.aov)
shapiro.test(x = dropOct7.resid) ## p>0.05
##### test homogeniety of variance
Oct7Resid <- ggplot(Oct7Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values", 
       subtitle = "Residual Plot-Oct. 7")+
  theme_jl
leveneTest(perDropOct7 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropOct7.aov)
#### Oct 13----
dropOct13.aov <- aov(perDropOct13 ~ treatment, data = drops)
summary.aov(dropOct13.aov)
Oct13Model <- data.frame(fit = fitted(dropOct13.aov),
                         res = resid(dropOct13.aov),
                         treat = drops$treatment)
#### test normality
Oct13QQ <- ggplot(drops, aes(sample = perDropOct13)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Oct. 13") +
  theme_jl
dropOct13.resid <- residuals(object = dropOct13.aov)
shapiro.test(x = dropOct13.resid) ## p>0.05
##### test homogeniety of variance
Oct13Resid <- ggplot(Oct13Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Oct. 13")+
  theme_jl
leveneTest(perDropOct13 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropOct13.aov)
#### Oct 21----
dropOct21.aov <- aov(perDropOct21 ~ treatment, data = drops)
summary.aov(dropOct21.aov)
Oct21Model <- data.frame(fit = fitted(dropOct21.aov),
                         res = resid(dropOct21.aov),
                         treat = drops$treatment)
#### test normality
Oct21QQ <- ggplot(drops, aes(sample = perDropOct21)) + 
  stat_qq() +
  stat_qq_line() +
  labs(y = "Sample",
       x = "Theoretical Quantiles",
       subtitle = "QQ Plot-Oct. 21") +
  theme_jl
dropOct21.resid <- residuals(object = dropOct21.aov)
shapiro.test(x = dropOct21.resid) ## p>0.05
##### test homogeniety of variance
Oct21Resid <- ggplot(Oct21Model, aes(x = fit, y = res))+
  geom_point()+
  geom_hline(yintercept = 0, colour = "pink")+
  scale_color_brewer(palette = "RdYlBu")+
  labs(y = "Residuals",
       x = "Fitted Values",
       subtitle = "Residual Plot-Oct. 21")+
  theme_jl
leveneTest(perDropOct21 ~ treatment, data = drops) ## p > 0.05
#### mean separation
TukeyHSD(dropOct21.aov)
## plot data----
treatmentCat <- c("ctrl", "avg", "naa", "ethephon")
sep <- c(rep("a", times = 20), rep(c("a", "b", "b", "a"), times = 4))
dropsLong %>%
  group_by(date, treatment) %>%
  summarise(avg.drop = mean(drops)) %>%
  arrange(date, match(treatment, treatmentCat)) %>%
  add_column(sep = sep) %>%
  ggplot(aes(x = date, y = avg.drop, color = treatment)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = mdy("8/26/2020"), colour = "pink", linetype = "dashed") +
  annotate("text", x = mdy("8/26/2020") - 1, y = 30, label = "Treatment Application", angle = 90, size = 3) +
  geom_text_repel(aes(x = date, y= avg.drop, label = c(rep(NA, times = 16), "a", "a", "a", "a",
                                                       rep(c("a", "b", "b", "a"), times = 4))), vjust=1, color="black",
            position = position_dodge(2), size= 3,parse = T, min.segment.length = 1.9) +
  scale_color_brewer(palette = "RdBu", labels = c("AVG", "Control", "Ethephon", "NAA")) +
  labs(y = "Cumulative Percent Drop (%)",
       x = "Date",
       color = "Treatment",
       title = "Cumulative Fruit Drop")+
  theme_jl
```

## Introduction

This project analyzes fruit drop count data for a Preharvest Drop study conducted from Aug.-Oct. 2020. Four treatments were applied on Aug. 26th: AVG, NAA, Ethephon, and an untreated control. Drop counts were done by counting the fruit that had dropped beneath the center tree of three tree plots. The plots were raked clean beneath the tree on Aug. 19th and then weekly drop counts were done on the following dates:

* August 25

* September 1

* September 8

* September 15

* September 22

* September 29

* October 7

* October 13

* October 21

After the final drop count, on Oct. 21, the remaining fruit on the tree were harvested to determine percent of fruit that dropped cummulatively throughout the experiment. 

## Calculate Percent Fruit Drop 

After loading a file for each drop count date and the tree count date, each drop date was joined into the tree counts...

```{r, eval=FALSE, warning=FALSE}
treeCounts %>%
  rowwise() %>%
  mutate(remainingFruit = sum(count1,count2,count3,count4,count5,count6,count7, na.rm = TRUE)) %>%
  select(row, tree, plot, treatment, rep, flag, remainingFruit) %>%
  left_join(dropAug25) %>%
  rename(Aug25Drop = dropped) %>%
```

Each date after Aug. 25 was successively added to determine cumulative dropped fruit from each tree...

```{r, eval=FALSE, warning=FALSE}
left_join(dropSep1) %>%
  rename(Sep1Drop = dropped)%>%
  mutate(cumDropSep1 = Aug25Drop + Sep1Drop) %>%
  left_join(dropSep8) %>%
  rename(Sep8Drop = dropped)%>%
  mutate(cumDropSep8 = Sep8Drop + cumDropSep1)
```

Finally, cumulative fruit drop was determined on each date was determined as a percentage of total fruit that was calculated...

```{r, eval=FALSE, warning=FALSE}
mutate(totalFruit = cumDropOct21 + remainingFruit) %>%
  mutate(perDropAug25 = ((Aug25Drop / totalFruit)*100),
         perDropSep1 = ((cumDropSep1 / totalFruit)*100),
         perDropSep8 = ((cumDropSep8 / totalFruit)*100),
         perDropSep15 = ((cumDropSep15 / totalFruit)*100),
         perDropSep22 = ((cumDropSep22 / totalFruit)*100),
         perDropSep29 = ((cumDropSep29 / totalFruit)*100),
         perDropOct7 = ((cumDropOct7 / totalFruit)*100),
         perDropOct13 = ((cumDropOct13 / totalFruit)*100),
         perDropOct21 = ((cumDropOct21 / totalFruit)*100))%>%
  select(row, tree, plot, treatment, rep, perDropAug25, perDropSep1, perDropSep8, perDropSep15, perDropSep22,
         perDropSep29, perDropOct7, perDropOct13, perDropOct21)-> drops
```

From there a long form data frame was made with plot info, percent fruit drop and date measured...

```{r, eval=FALSE, warning=FALSE}
drops %>%
  gather('perDropAug25', 'perDropSep1', 'perDropSep8', 'perDropSep15', 'perDropSep22', 'perDropSep29',
         'perDropOct7', 'perDropOct13', 'perDropOct21', key = "measDate", value = "drops") %>%
  mutate(date = case_when(measDate == 'perDropAug25' ~ mdy("8/25/2020"), ### add date
                   measDate == 'perDropSep1' ~ mdy("9/1/2020"),
                   measDate == 'perDropSep8' ~ mdy("9/8/2020"),
                   measDate == 'perDropSep15' ~ mdy("9/15/2020"),
                   measDate == 'perDropSep22' ~ mdy("9/22/2020"),
                   measDate == 'perDropSep29' ~ mdy("9/29/2020"),
                   measDate == 'perDropOct7' ~ mdy("10/7/2020"),
                   measDate == 'perDropOct13' ~ mdy("10/13/2020"),
                   measDate == 'perDropOct21' ~ mdy("10/21/2020"))) %>%
  select(row, tree, plot, treatment, rep, drops, date) -> dropsLong
```

## Explore Data

A box plot was created for each date, looked as though we don't see any treatment effect until Sep. 15th...

```{r, echo=FALSE, warning=FALSE, fig.width= 10, fig.height=10}
ggplot(dropsLong, aes(x = treatment, y = drops, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~ date, scales = "free") +
  scale_fill_brewer(palette = "RdYlBu") +
  labs(y = "Percent Fruit Drop (%)",
       x = "Treatment") +
  theme_jl
```

## ANOVA Assumptions

ANOVA for the full dataset showed that the interaction between date and drops was significant, so all further analysis was done within each date...

```{r, warning=FALSE}
drops.aov <- aov(drops ~ treatment*date, data = dropsLong)
summary.aov(drops.aov)
```

To check ANOVA assumptions an ANOVA model was built on each date...

```{r, eval=FALSE, warning=FALSE}
dropSep1.aov <- aov(perDropSep1 ~ treatment, data = drops)
```

Shapiro Wilkes test was done on the residuals to determine whether the data was normally distributed...

```{r, eval=FALSE, warning=FALSE}
dropSep1.resid <- residuals(object = dropSep1.aov)
shapiro.test(x = dropSep1.resid)
```

Then Levene's test was done to check for homogeneity of variance...

```{r, eval=FALSE, warning=FALSE}
leveneTest(perDropSep1 ~ treatment, data = drops)
```

The following are the p-values for these test statics on each date...

### **Aug. 25**

```{r, warning=FALSE}
shapiro.test(x = dropAug25.resid)
```

p<0.05 and therefore data was transformed via square root...

```{r, warning=FALSE}
shapiro.test(x = sqrt(dropAug25.resid))
```

The data was now normally distributed and levene's test was not significant...

```{r, warning=FALSE}
leveneTest(perDropAug25 ~ treatment, data = drops)
```

### **Sep. 1**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropSep1.resid)
```

```{r, warning=FALSE}
leveneTest(perDropSep1 ~ treatment, data = drops)
```

### **Sep. 8**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropSep8.resid)
```

```{r, warning=FALSE}
leveneTest(perDropSep8 ~ treatment, data = drops)
```

### **Sep. 15**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropSep15.resid)
```

```{r, warning=FALSE}
leveneTest(perDropSep15 ~ treatment, data = drops)
```

### **Sep. 22**

Data was not normally distributed...

```{r, warning=FALSE}
shapiro.test(x = dropSep22.resid)
```

Taking the negative square root normalized the data...

```{r, warning=FALSE}
shapiro.test(x = (dropSep22.resid)^(-.5))
```

Variances were equal among groups...

```{r, warning=FALSE}
leveneTest(perDropSep22 ~ treatment, data = drops)
```

### **Sep. 29**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropSep29.resid)
```

```{r, warning=FALSE}
leveneTest(perDropSep29 ~ treatment, data = drops)
```

### **Oct. 7**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropOct7.resid)
```

```{r, warning=FALSE}
leveneTest(perDropOct7 ~ treatment, data = drops)
```

### **Oct. 13**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropOct13.resid)
```

```{r, warning=FALSE}
leveneTest(perDropOct13 ~ treatment, data = drops)
```

### **Oct. 21**

p<0.05 for both test statistics...

```{r, warning=FALSE}
shapiro.test(x = dropOct21.resid)
```

```{r, warning=FALSE}
leveneTest(perDropOct21 ~ treatment, data = drops)
```

Following are diagnostic plots for the residuals to assess homogeneity of variance and a QQ plot for normality:
```{r, echo=FALSE, warning=FALSE, fig.width= 10, fig.height=10}
ggarrange(Aug25QQ, Aug25Resid, Sep1QQ, Sep1Resid, Sep8QQ, Sep8Resid, ncol = 2, nrow = 3)
ggarrange(Sep15QQ, Sep15Resid, Sep22QQ, Sep22Resid, Sep29QQ, Sep29Resid, ncol = 2, nrow = 3)
ggarrange(Oct7QQ, Oct7Resid, Oct13QQ, Oct13Resid, Oct21QQ, Oct21Resid, ncol = 2, nrow = 3)
```

## Mean Separation

After transformation of drop data on Aug. 25 and Sep. 22 mean separation was done from the ANOVA model using Tukey's HSD, below is the resulting p-values for each pairwise comparison on each measurement date...

### **Aug. 25**

```{r, warning=FALSE}
dropAug25.norm.aov <- aov(sqrt(perDropAug25) ~ treatment, data = drops) 
TukeyHSD(dropAug25.norm.aov)
```

### **Sep. 1**

```{r, warning=FALSE}
TukeyHSD(dropSep1.aov)
```

### **Sep. 8**

```{r, warning=FALSE}
TukeyHSD(dropSep8.aov)
```

### **Sep. 15**

```{r, warning=FALSE}
TukeyHSD(dropSep15.aov)
```

### **Sep. 22**

```{r, warning=FALSE}
dropSep22.norm.aov <- aov((perDropSep22)^(-.5) ~ treatment, data = drops)
TukeyHSD(dropSep22.norm.aov)
```

### **Sep. 29**

```{r, warning=FALSE}
TukeyHSD(dropSep29.aov)
```

### **Oct. 7**

```{r, warning=FALSE}
TukeyHSD(dropOct7.aov)
```

### **Oct. 13**

```{r, warning=FALSE}
TukeyHSD(dropOct13.aov)
```

### **Oct. 21**

```{r, warning=FALSE}
TukeyHSD(dropOct21.aov)
```

Below is the resulting figure...

```{r, echo=FALSE, warning=FALSE}
treatmentCat <- c("ctrl", "avg", "naa", "ethephon")
sep <- c(rep("a", times = 20), rep(c("a", "b", "b", "a"), times = 4))
dropsLong %>%
  group_by(date, treatment) %>%
  summarise(avg.drop = mean(drops)) %>%
  arrange(date, match(treatment, treatmentCat)) %>%
  add_column(sep = sep) %>%
  ggplot(aes(x = date, y = avg.drop, color = treatment)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = mdy("8/26/2020"), colour = "pink", linetype = "dashed") +
  annotate("text", x = mdy("8/26/2020") - 1, y = 30, label = "Treatment Application", angle = 90, size = 3) +
  geom_text_repel(aes(x = date, y= avg.drop, label = c(rep(NA, times = 16), "a", "a", "a", "a",
                                                       rep(c("a", "b", "b", "a"), times = 4))), vjust=1, color="black",
            position = position_dodge(2), size= 3,parse = T, min.segment.length = 1.9) +
  scale_color_brewer(palette = "RdBu", labels = c("AVG", "Control", "Ethephon", "NAA")) +
  labs(y = "Cumulative Percent Drop (%)",
       x = "Date",
       color = "Treatment",
       title = "Cumulative Fruit Drop")+
  theme_jl
```