---
title: "Fruit Sorter Data Clean Up"
author: "Jimmy Larson"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(car)
library(lubridate)
library(RColorBrewer)
library(ggrepel)
library(hrbrthemes)
library(latex2exp)
```


## Introduction

This is file describes the script that takes cleans data from a fruit sorting line. Near infrared spectroscopy and cameras are used to calculate fruit size quality parameters. The sorter provides a .csv file for each experimental unit that runs through the sorter line. All files are placed into a folder. This script takes each of those files in the folder, selects fruit weight, length, diameter, blush percentage, and apple quality; then binds each file together.

## User Inputs

The user puts the folder info into the path, sets what they want to save the file as, a determines the pattern from the file name to add to the data frame. 

```{r, message=FALSE}
path <- "data/2020/sorter/20PFD_H.-2/" # enter path here
saveFile <- "sorter_data_aug28.csv" # enter info to save file as
pattern <- "_\\w+_...." # set pattern to extract treatment and rep info from file name
```

## For Loop 

Before beginning the for loop, a list of file names and the number of files were set as files...

```{r, message=FALSE}
fileName <- list.files(path= path, pattern="*.csv", full.names=TRUE, recursive=FALSE)
nFiles <- length(fileName)
```

An output file was then initiated...

```{r, message=FALSE}
out.file <- ""
```

A for loop was then created...

```{r, message=FALSE, warning=FALSE}
for (i in 1:nFiles) {
  ## load in data
  file <- read_delim(fileName[i], ";", escape_double = FALSE, trim_ws = TRUE)
  ## select rows----
  file <- file[grep(">|<", file$ClassName),]
  ## extract treatment info from filename----
  string <- fileName[i]
  trt <- regmatches(string, regexpr(pattern, string))
  ## add treatment info to file----
  file %>%
    mutate(treatment_rep = trt) %>%
    separate(treatment_rep, sep = "_", into = c("blank","treatment", "rep")) %>%
    select(treatment, rep, ClassName, Weight, BlushPercentage, Diameter, AppleQuality, Length) -> file
  ## bind files together
  out.file <- rbind(out.file, file)
}
out.file <- out.file[-1,]
```

Here is the resulting data frame...

```{R, message = FALSE}
head(out.file)
```

## Steps in For Loop

Each file was read in...

```{r, eval=FALSE}
file <- read_delim(fileName[i], ";", escape_double = FALSE, trim_ws = TRUE)
```

The important data was selected from the rest of the file...

```{r, eval=FALSE}
file <- file[grep(">|<", file$ClassName),]
```

Treatment info was extracted from the file name...

```{r, eval=FALSE}
  string <- fileName[i]
  trt <- regmatches(string, regexpr(pattern, string))
```

Treatment info was then added into the dataframe and columns were selected...

```{r, eval=FALSE}
  file %>%
    mutate(treatment_rep = trt) %>%
    separate(treatment_rep, sep = "_", into = c("blank","treatment", "rep")) %>%
    select(treatment, rep, ClassName, Weight, BlushPercentage, Diameter, AppleQuality, Length) -> file
```

Files are then binded together in output file...

```{r, eval=FALSE}
out.file <- rbind(out.file, file)
```

## Output File

The top blank line that was used to initiate the output file was deleted and the full data frame was saved...

```{r, eval=FALSE}
out.file <- out.file[-1,]
write.csv(out.file, saveFile)
```